Further Modeling of $q_{95}$ Windows for the Suppression of Edge Localized Modes by Resonant Magnetic Perturbations in the DIII-D Tokamak

### 2020-10-31 Papers/Paper2/Paper.tex
Lastest version

``Asymptotic matching'' is by far the most efficient method for modeling the response of a tokamak plasma to an RMP.\cite{fkr,rfa, am1,am2,am3} 
According to the asymptotic matching approach, the response of the plasma to the applied RMP is governed by a combination of flux-freezing and
perturbed force balance---this combination is usually referred to as ``marginally-stable ideal-magnetohydrodynamics (MHD)"---everywhere in the plasma apart from a number of relatively narrow (in the radial direction) regions in which the applied
perturbation resonates with the equilibrium magnetic field. Magnetic reconnection can take place within the resonant regions to
produce relatively narrow magnetic island chains. Within the resonant regions, the plasma response is governed by  linear or nonlinear 
 two-fluid resistive-MHD, depending on whether the induced magnetic island widths are smaller or larger, respectively, than the corresponding linear layer widths. 
 Thus, when employing the asymptotic matching approach, the equations of marginally-stable ideal-MHD
are solved in the so-called ``outer region'' that comprises most of the plasma (and the surrounding vacuum), the equations of 
linear/nonlinear two-fluid resistive-MHD are solved in the various resonant layers that constitute the so-called ``inner region'', and the two
sets of solutions are then asymptotically matched to one another.

@@

\subsection{Linear Response Regime}
In the linear response regime, the reconnected magnetic flux induced by the RMP at the
rational surface is governed by\,\cite{rfa,rfb,rfc}
\begin{equation}\label{e1}
\frac{\delta_s}{r_s}\,\tau_R\left(\frac{d}{dt}+{\rm i}\,\omega_s\right){\mit\Psi}_s= E_{ss}\,{\mit\Psi}_s+E_{sv}\,{\mit\Psi}_v,
\end{equation}
where $\delta_s$ is the linear layer width, $\tau_R=\mu_0\,r_s^{\,2}\,\sigma(r_s)$ the global resistive diffusion timescale,
 and
\begin{equation}
\omega_s (t)=m\,{\mit\Omega}_\theta(r_s,t)-n\,{\mit\Omega}_\phi(r_s,t).
\end{equation}
Here, $\sigma(r)$, ${\mit\Omega}_\theta(r,t)$, and ${\mit\Omega}_\phi(r,t)$ are the 
plasma conductivity, poloidal angular velocity, and toroidal angular velocity profiles, respectively. 
Note that, in this paper, we are assuming that the constant-$\psi$ approximation\,\cite{fkr} holds at the rational surface. (It is demonstrated in
Ref.~\onlinecite{rf1} that the appropriate linear response regime for an RMP resonant in the pedestal of a typical DIII-D H-mode discharge
is the so-called  ``first semi-collisional regime'', which is indeed a constant-$\psi$ response regime. Moreover, if the appropriate linear response regime
is constant-$\psi$ then so is the appropriate nonlinear response regime, because the constant-$\psi$ constraint is
easier to satisfy in the nonlinear regime than in the linear regime.\cite{rfb,rffig})

@@

The upper panel of Fig.~\ref{fig3} shows the widths of the various vacuum magnetic island chains driven by the applied $n=3$ RMP in the pedestal region as functions of $
\overline{q_{95}}$.
The vacuum island widths are the  expected island widths in the absence of any shielding of driven magnetic reconnection due to
plasma flow; in other words, they are the  widths of the island chains that would be driven if all of the natural frequencies were zero. It can be seen that
the driven magnetic island chains extend over most of the pedestal, implying a very significant degradation in pedestal energy and particle confinement due to the
expected flattening of the density and temperature profiles across the chains.\cite{flat} Moreover, the driven island chains at the
bottom of the pedestal overlap, implying that the magnetic field in this region is rendered stochastic. The lower panel
of Fig.~\ref{fig3} shows the actual widths of the various magnetic island chains driven by the applied $n=3$ RMP as functions of time. It is apparent 
that the true island widths are generally much smaller than the vacuum island widths, as a consequence of the shielding of driven magnetic
reconnection due to plasma flow. However, the shielding breaks down at various rational surfaces when the associated natural frequency is zero (see the lower panel of
Fig.~\ref{fig3}). In this situation, the driven island width is similar to the vacuum island width.

@@

Finally, the best agreement between theory and experiment is obtained by assuming that, in the absence of the RMP, the 
 natural frequency of the tearing mode resonant at the $k$th rational surface is given by\,\cite{rftor1}
\begin{equation}\label{eeb}
\varpi_{k\,0} = \varpi_{{\rm eb}\,k}\equiv -n\,(\omega_E)_{{\mit\Psi}_{p\,k}}.
\end{equation}

@@

@@

@@

\bibitem{rftor1} R.~Fitzpatrick,  ``Modeling $q_{95}$ Windows for the Suppression of Edge Localized
Modes by Resonant Magnetic Perturbations in the DIII-D Tokamak", submitted to Physics of Plasmas (2020).

@@

@@

\begin{figure}
\includegraphics[height=6in]{fig3.pdf}
\caption{Composite linear/nonlinear calculation with the no-slip constraint imposed at all rational surfaces, and the natural frequencies
in the absence of the RMP determined by the local  ${\bf E}\times {\bf B}$
frequency. Top Panel: Full  $n=3$ vacuum island widths as functions of the least-squares linear fit to $q_{95}$ versus time 
in   DIII-D discharge \#145380.
Bottom Panel:  Full $n=3$ island widths as functions of time
in   DIII-D discharge \#145380. The blue, yellow, cyan, magenta, brown, pink,
purple, and orange  areas correspond to $m=7$, 8, 9, 10, 11, 12, 13, and 14, respectively. The yellow vertical bands indicate the ELM-suppression/mitigation windows. 
The horizontal dotted lines indicate the top of the pedestal, ${\mit\Psi}_N=0.925$.} \label{fig3}
\end{figure}

@@

\begin{figure}
\includegraphics[height=6in]{fig5.pdf}
\caption{Composite linear/nonlinear calculation with the no-slip constraint relaxed at all rational surfaces, and the natural frequencies
in the absence of the RMP determined by the local  ${\bf E}\times {\bf B}$
frequency. Top Panel: Full  $n=3$ vacuum island widths as functions of the least-squares linear fit to $q_{95}$ versus time 
in   DIII-D discharge \#145380.
Bottom Panel:  Full $n=3$ island widths as functions of time
in   DIII-D discharge \#145380. The blue, yellow, cyan, magenta, brown, pink,
purple, and orange  areas correspond to $m=7$, 8, 9, 10, 11, 12, 13, and 14, respectively. The yellow vertical bands indicate the ELM-suppression/mitigation windows. 
The horizontal dotted lines indicate the top of the pedestal, ${\mit\Psi}_N=0.925$.} \label{fig5}
\end{figure}

@@

\begin{figure}
\includegraphics[height=6in]{fig8.pdf}
\caption{Linear calculation with the no-slip constraint relaxed at all rational surfaces, and  the natural frequencies
in the absence of the RMP determined by linear theory. Top Panel: Full  $n=3$ vacuum island widths as functions of the least-squares linear fit to $q_{95}$ versus time 
in   DIII-D discharge \#145380.
Bottom Panel:  Full $n=3$ island widths as functions of time
in   DIII-D discharge \#145380. The blue, yellow, cyan, magenta, brown, pink,
purple, and orange  areas correspond to $m=7$, 8, 9, 10, 11, 12, 13, and 14, respectively. The yellow vertical bands indicate the ELM-suppression/mitigation windows. 
The horizontal dotted lines indicate the top of the pedestal, ${\mit\Psi}_N=0.925$.} \label{fig8}
\end{figure}

@@

\begin{figure}
\includegraphics[height=6in]{fig10.pdf}
\caption{Composite linear/nonlinear calculation with the no-slip constraint relaxed at all rational surfaces, and  the natural frequencies
in the absence of the RMP determined by composite linear/nonlinear theory. Top Panel: Full  $n=3$ vacuum island widths as functions of the least-squares linear fit to $q_{95}$ versus time 
in   DIII-D discharge \#145380.
Bottom Panel:  Full $n=3$ island widths as functions of time
in   DIII-D discharge \#145380. The blue, yellow, cyan, magenta, brown, pink,
purple, and orange  areas correspond to $m=7$, 8, 9, 10, 11, 12, 13, and 14, respectively. The yellow vertical bands indicate the ELM-suppression/mitigation windows. 
The horizontal dotted lines indicate the top of the pedestal, ${\mit\Psi}_N=0.925$.} \label{fig10}
\end{figure}
\end{document}

>>>>

``Asymptotic matching'' is by far the most efficient method for modeling the response of a tokamak plasma to an RMP.\cite{fkr,coppi,ruth,ara,pletz,rfa, am1,tokuda,brennan,am2,am3} 
According to the asymptotic matching approach, the response of the plasma to the applied RMP is governed by a combination of flux-freezing and
perturbed force balance---this combination is usually referred to as ``marginally-stable ideal-magnetohydrodynamics (MHD)"---everywhere in the plasma apart from a number of relatively narrow (in the radial direction) regions in which the applied
perturbation resonates with the equilibrium magnetic field. Magnetic reconnection can take place within the resonant regions to
produce relatively narrow magnetic island chains. Within the resonant regions, the plasma response is governed by  linear or nonlinear 
 two-fluid resistive-MHD, depending on whether the induced magnetic island widths are smaller or larger, respectively, than the corresponding linear layer widths. 
 Thus, when employing the asymptotic matching approach, the equations of marginally-stable ideal-MHD
are solved in the so-called ``outer region'' that comprises most of the plasma (and the surrounding vacuum), the equations of 
linear/nonlinear two-fluid resistive-MHD are solved in the various resonant layers that constitute the so-called ``inner region'', and the two
sets of solutions are then asymptotically matched to one another.

@@

\subsection{Linear Response Regime}
In the linear response regime, the reconnected magnetic flux induced by the RMP at the
rational surface is governed by\,\cite{rfa,rfb,rfc}
\begin{equation}\label{e1}
\frac{\delta_s}{r_s}\,\tau_R\left(\frac{d}{dt}+{\rm i}\,\omega_s\right){\mit\Psi}_s= E_{ss}\,{\mit\Psi}_s+E_{sv}\,{\mit\Psi}_v,
\end{equation}
where $\delta_s$ is the linear layer width, $\tau_R=\mu_0\,r_s^{\,2}\,\sigma(r_s)$ the global resistive diffusion timescale,
 and
\begin{equation}
\omega_s (t)=m\,{\mit\Omega}_\theta(r_s,t)-n\,{\mit\Omega}_\phi(r_s,t).
\end{equation}
Here, $\sigma(r)$, ${\mit\Omega}_\theta(r,t)$, and ${\mit\Omega}_\phi(r,t)$ are the 
plasma electrical conductivity, poloidal angular velocity, and toroidal angular velocity profiles, respectively. 
Note that, in this paper, we are assuming that the constant-$\psi$ approximation\,\cite{fkr} holds at the rational surface. (It is demonstrated in
Ref.~\onlinecite{rf1} that the appropriate linear response regime for an RMP resonant in the pedestal of a typical DIII-D H-mode discharge
is the so-called  ``first semi-collisional regime'', which is indeed a constant-$\psi$ response regime. Moreover, if the appropriate linear response regime
is constant-$\psi$ then so is the appropriate nonlinear response regime, because the constant-$\psi$ constraint is
easier to satisfy in the nonlinear regime than in the linear regime.\cite{rfb,rffig})

@@

The upper panel of Fig.~\ref{fig3} shows the widths of the various vacuum magnetic island chains driven by the applied $n=3$ RMP in the pedestal region as functions of $
\overline{q_{95}}$.
The vacuum island widths are the  expected island widths in the absence of any shielding of driven magnetic reconnection due to
plasma flow; in other words, they are the  widths of the island chains that would be driven if all of the natural frequencies were zero. It can be seen that
the driven magnetic island chains extend over most of the pedestal, implying a very significant degradation in pedestal energy and particle confinement due to the
expected flattening of the density and temperature profiles across the chains.\cite{flat} Moreover, the driven island chains at the
bottom of the pedestal overlap, implying that the magnetic field in this region is rendered stochastic. The lower panel
of Fig.~\ref{fig3} shows the predicted widths of the various magnetic island chains driven by the applied $n=3$ RMP as functions of time. It is apparent 
that the predicted island widths are generally much smaller than the vacuum island widths, as a consequence of the shielding of driven magnetic
reconnection due to plasma flow. However, the shielding breaks down at various rational surfaces when the associated natural frequency is zero (see the lower panel of
Fig.~\ref{fig3}). In this situation, the driven island width is similar to the vacuum island width.

@@

Finally, the best agreement between theory and experiment is obtained by assuming that, in the absence of the RMP, the 
 natural frequency of the tearing mode resonant at the $k$th rational surface is given by\,\cite{rftor1,heyn}
\begin{equation}\label{eeb}
\varpi_{k\,0} = \varpi_{{\rm eb}\,k}\equiv -n\,(\omega_E)_{{\mit\Psi}_{p\,k}}.
\end{equation}

@@

\bibitem{coppi} B.~Coppi, J.M.~Greene, and J.L.~Johnson, Nucl.\ Fusion {\bf 6}, 101 (1966).

\bibitem{pletz} A.~Pletzer, and R.L.~Dewar, J.\ Plasma Physics {\bf 45}, 427 (1991).

@@

\bibitem{tokuda} S.~Tokuda, Nucl.\ Fusion {\bf 41}, 1037 (2001).

\bibitem{brennan} D.P.~Brennan, R.J.~La Haye, A.D.~Turnbull, M.S.~Chu, T.H.~Jensen, L.L.~Lao, T.C.~Luce, P.A.~Politzer, and E.J.~Strait,
Phys.\ Plasmas {\bf 10}, 1643 (2003).

@@

\bibitem{rftor1} R.~Fitzpatrick,  Phys.\ Plasmas {\bf 27}, 102511 (2020).

@@

\bibitem{heyn} M.F.~Heyn, I.B.~Ivanov, S.V.~Kasilov, W.~Kernbichler, I.~Joseph, R.A.~Moyer,  and A.M.~Runov, Nucl.\ Fusion {\bf 48}, 024005 (2008).

@@

\begin{figure}
\includegraphics[height=6in]{fig3.pdf}
\caption{Composite linear/nonlinear calculation with the no-slip constraint imposed at all rational surfaces, and the natural frequencies
in the absence of the RMP determined by the local  ${\bf E}\times {\bf B}$
frequency. Top Panel: Full  $n=3$ vacuum island widths as functions of the least-squares linear fit to $q_{95}$ versus time 
in   DIII-D discharge \#145380.
Bottom Panel:  Full $n=3$ island widths as functions of time
in   DIII-D discharge \#145380. The yellow, cyan, magenta, brown, pink,
purple, and orange  areas correspond to $m=8$, 9, 10, 11, 12, 13, and 14, respectively. The yellow vertical bands indicate the ELM-suppression/mitigation windows. 
The horizontal dotted lines indicate the top of the pedestal, ${\mit\Psi}_N=0.925$.} \label{fig3}
\end{figure}

@@

\begin{figure}
\includegraphics[height=6in]{fig5.pdf}
\caption{Composite linear/nonlinear calculation with the no-slip constraint relaxed at all rational surfaces, and the natural frequencies
in the absence of the RMP determined by the local  ${\bf E}\times {\bf B}$
frequency. Top Panel: Full  $n=3$ vacuum island widths as functions of the least-squares linear fit to $q_{95}$ versus time 
in   DIII-D discharge \#145380.
Bottom Panel:  Full $n=3$ island widths as functions of time
in   DIII-D discharge \#145380. The yellow, cyan, magenta, brown, pink,
purple, and orange  areas correspond to $m=8$, 9, 10, 11, 12, 13, and 14, respectively. The yellow vertical bands indicate the ELM-suppression/mitigation windows. 
The horizontal dotted lines indicate the top of the pedestal, ${\mit\Psi}_N=0.925$.} \label{fig5}
\end{figure}

@@

\begin{figure}
\includegraphics[height=6in]{fig8.pdf}
\caption{Linear calculation with the no-slip constraint relaxed at all rational surfaces, and  the natural frequencies
in the absence of the RMP determined by linear theory. Top Panel: Full  $n=3$ vacuum island widths as functions of the least-squares linear fit to $q_{95}$ versus time 
in   DIII-D discharge \#145380.
Bottom Panel:  Full $n=3$ island widths as functions of time
in   DIII-D discharge \#145380. The yellow, cyan, magenta, brown, pink,
purple, and orange  areas correspond to $m=8$, 9, 10, 11, 12, 13, and 14, respectively. The yellow vertical bands indicate the ELM-suppression/mitigation windows. 
The horizontal dotted lines indicate the top of the pedestal, ${\mit\Psi}_N=0.925$.} \label{fig8}
\end{figure}

@@

\begin{figure}
\includegraphics[height=6in]{fig10.pdf}
\caption{Composite linear/nonlinear calculation with the no-slip constraint relaxed at all rational surfaces, and  the natural frequencies
in the absence of the RMP determined by composite linear/nonlinear theory. Top Panel: Full  $n=3$ vacuum island widths as functions of the least-squares linear fit to $q_{95}$ versus time 
in   DIII-D discharge \#145380.
Bottom Panel:  Full $n=3$ island widths as functions of time
in   DIII-D discharge \#145380. The yellow, cyan, magenta, brown, pink,
purple, and orange  areas correspond to $m=8$, 9, 10, 11, 12, 13, and 14, respectively. The yellow vertical bands indicate the ELM-suppression/mitigation windows. 
The horizontal dotted lines indicate the top of the pedestal, ${\mit\Psi}_N=0.925$.} \label{fig10}
\end{figure}
\end{document}
